---
title: "Attribute analysis - Distances"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, messages=FALSE, results='hide'}
#garbage collection + clear RAM
rm(list = ls(all.name = TRUE))
gc()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include=FALSE}
  #load libraries
  library(knitr)
  library(ggplot2)
  library(scales)
  library(gridExtra)

```

## Connect to the database first
```{r echo=FALSE}
read_chunk('../../DBConnection/ConnectPostgres.R')
```

```{r connectDB, echo=FALSE}
```

##Get the tweets
```{r gettwitterfeeddata, cache=FALSE, echo=FALSE}
users <- dbGetQuery(con, "SELECT u.dist_latlon_vs_tz, u.dist_latlon_vs_loc, u.dist_tz_vs_loc, tz.continent from main.experiment_user u left join main.timezone_r tz on tz.timezone = u.timezone")
```

###Display latitude, longitude vs timezone
```{r ll_tz}
df <- data.frame(users$dist_latlon_vs_tz, users$continent)
user.data <- df[complete.cases(df),]
colnames(user.data) = c("dist_latlon_vs_tz","continent")
p2 <- ggplot(user.data, aes(x=continent, y=dist_latlon_vs_tz)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4) +
  stat_summary(fun.y=mean, geom="point", shape=23, size=4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=1) 
  #geom_jitter(shape=16, position=position_jitter(0.2))
p2
```
###Display latitude, longitude vs location
```{r ll_loc}
df <- data.frame(users$dist_latlon_vs_loc, users$continent)
user.data <- df[complete.cases(df),]
colnames(user.data) = c("dist_latlon_vs_loc","continent")
p3 <- ggplot(user.data, aes(x=continent, y=dist_latlon_vs_loc)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4) +
  stat_summary(fun.y=mean, geom="point", shape=23, size=4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=1) 
  #geom_jitter(shape=16, position=position_jitter(0.2))
p3
```
###Display location vs timezone
```{r loc_tz}
df <- data.frame(users$dist_tz_vs_loc, users$continent)
user.data <- df[complete.cases(df),]
colnames(user.data) = c("dist_tz_vs_loc","continent")
p4 <- ggplot(user.data, aes(x=continent, y=dist_tz_vs_loc)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4) +
  stat_summary(fun.y=mean, geom="point", shape=23, size=4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=1) 
  #geom_jitter(shape=16, position=position_jitter(0.2))
p4
```

##All in one grid
```{r grid}
grid.arrange(p2, p3, p4, nrow=1)
```

##Combined results
```{r loc_total}
df <- data.frame((users$dist_latlon_vs_tz+users$dist_latlon_vs_loc+users$dist_tz_vs_loc)/3, users$continent)
user.data <- df[complete.cases(df),]
colnames(user.data) = c("dist","continent")
ggplot(user.data, aes(x=continent, y=dist)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4) +
  stat_summary(fun.y=mean, geom="point", shape=23, size=4) 
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=1) 
  #geom_jitter(shape=16, position=position_jitter(0.2))
```

###Score the data

##first create a score
```{r score}
exp_no <- 1
period_no <- 1
factor_no <- 5

sql <- paste("DELETE FROM main.experiment_user_score where factor_no = 5",sep="")
dbSendQuery(con, sql)

sql <- paste("INSERT INTO main.experiment_user_score(experiment_no, period_no, userid, factor_no, idi_full)",sep="")
sql <- paste(sql, "select experiment_no, period_no, userid, 5,", sep="")
sql <- paste(sql, "sum((coalesce(dist_latlon_vs_tz,0)+coalesce(dist_latlon_vs_loc,0)+coalesce(dist_tz_vs_loc,0)))/sum(case when dist_latlon_vs_tz is null and dist_latlon_vs_loc is null and dist_tz_vs_loc is null then 1 else 0 end + case when dist_latlon_vs_tz is not null then 1 else 0 end + case when dist_latlon_vs_loc is not null then 1 else 0 end + case when dist_tz_vs_loc is not null then 1 else 0 end)", sep="")    
sql <- paste(sql, " from main.experiment_user", sep="")
sql <- paste(sql, " group by experiment_no, period_no, userid", sep="")
dbSendQuery(con, sql)
    
```

##normalise the score and show results
The standard formula to normalize data =(value-min)/(max-min) 
R has a built in function to do this
```{r score_n}
user.score <- dbGetQuery(con, "SELECT s.idi_full, tz.continent, s.userid from main.experiment_user_score s join main.experiment_user u on u.userid = s.userid left join main.timezone_r tz on tz.timezone = u.timezone where s.factor_no = 5 and s.experiment_no = u.experiment_no and s.period_no = u.period_no")

user.scaled_score <- data.frame(as.data.frame( scale(user.score[1] )), user.score[2], user.score[3])
colnames(user.scaled_score) = c("idi","continent","userid")

ggplot(user.scaled_score, aes(x=continent, y=idi)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4) +
  stat_summary(fun.y=mean, geom="point", shape=23, size=4) 

```


##Outlier detection
Use Tukey's method to update all scores that were outliers

```{r outlier, results='hide'}

markoutlier <- function(x, exp_no, period_no, factor_no){
  sql <- paste("update main.experiment_user_score set outlier_full=1",sep="")
  sql <- paste(sql," where userid='",x["userid"] ,"'",sep="")
  sql <- paste(sql, " and experiment_no=",exp_no,sep="")
  sql <- paste(sql, " and period_no=",period_no,sep="")
  sql <- paste(sql, " and factor_no=",factor_no,sep="")
  dbSendQuery(con, sql, echo=FALSE)  
}

#TODO outliers per continent
continents <- unique(user.scaled_score$continent)

  user.continent_score <- user.scaled_score
  outlier <- boxplot.stats(user.continent_score$idi, coef = 1.5)$out
  user.outlier <- user.continent_score[user.continent_score$idi %in% outlier,]
  apply(user.outlier, 1, markoutlier, exp_no=exp_no, period_no=period_no,factor_no=factor_no)
  #na1 <- nrow(user.outlier)
  #Outliers identified:
  #na1
  #Propotion (%) of outliers:
  #round(na1 / sum(!is.na(user.continent_score$idi))*100, 1)

```

Total outliers: `r nrow(user.outlier)` out of `r nrow(user.scaled_score)`


```{r closeDB}
```