---
title: "Multi variate analysis"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, messages=FALSE, results='hide'}
#garbage collection + clear RAM
rm(list = ls(all.name = TRUE))
gc()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include=FALSE}
  #load libraries
  library(knitr)
  library(ggplot2)
  library(scales)
```

Model based approaches assume a variety of data models and apply maximum likelihood estimation and Bayes criteria to identify the most likely model and number of clusters. Specifically, the Mclust( ) function in the mclust package selects the optimal model according to BIC for EM initialized by hierarchical clustering for parameterized Gaussian mixture models.

## Connect to the database first
```{r echo=FALSE}
read_chunk('../DBConnection/ConnectPostgres.R')
```

```{r connectDB}
```

##Get the tweets
```{r gettwitterfeeddata, cache=FALSE}
users <- dbGetQuery(con, "SELECT * from main.experiment_user_shortest")
```

Total amount of users in the corpus: `r users.total <- nrow(users)
  users.total`

```{r preview_users}
#Preview user dataset  
head(users)

#scale the dataset
df <- scale(users[,3:10])

#Preview scaled dataset
head(df)
```

##First investigate the multivariate data

### scatterplot
User a scatterplot to investigate the different dimensions of data
```{r cluster_scatter}
library(car)

scatterplotMatrix(df)
```

### profileplot
To find the variance within a dimention
```{r cluster_profile}
  makeProfilePlot <- function(mylist,names)
  {
     require(RColorBrewer)
     # find out how many variables we want to include
     numvariables <- length(mylist)
     # choose 'numvariables' random colours
     colours <- brewer.pal(numvariables,"Set1")
     # find out the minimum and maximum values of the variables:
     mymin <- 1e+20
     mymax <- 1e-20
     for (i in 1:numvariables)
     {
        vectori <- mylist[[i]]
        mini <- min(vectori)
        maxi <- max(vectori)
        if (mini < mymin) { mymin <- mini }
        if (maxi > mymax) { mymax <- maxi }
     }
     # plot the variables
     for (i in 1:numvariables)
     {
        vectori <- mylist[[i]]
        namei <- names[i]
        colouri <- colours[i]
        if (i == 1) { plot(vectori,col=colouri,type="l",ylim=c(mymin,mymax)) }
        else         { points(vectori, col=colouri,type="l")                                     }
        lastxval <- length(vectori)
        lastyval <- vectori[length(vectori)]
        text((lastxval-10),(lastyval),namei,col="black",cex=0.6)
     }
  }

  library(RColorBrewer)
  makeProfilePlot(df,colnames(df))
  
```
  
### correlations
Try to find the dimentions which are most correlated with each other
```{r cluster_correlate}
mosthighlycorrelated <- function(mydataframe,numtoreport)
{
   # find the correlations
   cormatrix <- cor(mydataframe)
   # set the correlations on the diagonal or lower triangle to zero,
   # so they will not be reported as the highest ones:
   diag(cormatrix) <- 0
   cormatrix[lower.tri(cormatrix)] <- 0
   # flatten the matrix into a dataframe for easy sorting
   fm <- as.data.frame(as.table(cormatrix))
   # assign human-friendly names
   names(fm) <- c("First.Variable", "Second.Variable","Correlation")
   # sort and print the top n correlations
   head(fm[order(abs(fm$Correlation),decreasing=T),],n=numtoreport)
}

mosthighlycorrelated(df, 10)

```


```{r closeDB}
```