---
title: "Tweet Content Information"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
  #load libraries
  library(RPostgreSQL)
```

## Connect to the database first
```{r}
#loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")
#creates a connection to the postgres database
#note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "twitter",
host = "localhost", port = 5432,
user = "postgres", password = "")

```

Connection success: `r dbExistsTable(con, c("main","experiment_tweets_shortest"))`

##Pull data from the database
We will only pull the data for one user to understand the structure and content initially

```{r}
tweets <- dbGetQuery(con, "SELECT * from main.experiment_tweets_shortest where \"USERNAME\" = 'Londs_'")
```

Amount of records retrieved: `r nrow(tweets)`

## Structure of the data

```{r}
str(tweets)
```

## Summary of the data

```{r}
summary(tweets)
```

## Find NA (not applicables) in dataset 
These could be a problem later in data analysis


