---
title: "Tweet Account Info"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, messages=FALSE, results='hide'}
#garbage collection + clear RAM
rm(list = ls(all.name = TRUE))
gc()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include=FALSE}
  #load libraries
  library(knitr)
  library(ggplot2)
  library(scales)
```

In this section we will explore the user data we have available.
The user data is an aggregation of the information available in the tweets themselves, i.e. only the data pertaining to the account and not the actual tweet.
My aggregating this data we get a view of the account itself.

## Connect to the database first
```{r echo=FALSE}
read_chunk('../DBConnection/ConnectPostgres.R')
```

```{r connectDB}
```

##Get the users
```{r gettwitterfeeddata, cache=FALSE}
users <- dbGetQuery(con, "SELECT * from main.experiment_user_shortest")
```

Total amount of users in the corpus: `r users.total <- nrow(users)
  users.total`

```{r preview_users}
#Preview user dataset  
head(users)

dim(users)

str(users)

summary(users)


#scale the dataset
df <- scale(users[,3:10])

#Preview scaled dataset
head(df)
```



```{r closeDB}
```