---
title: "Tweet Corpus Information"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
```

```{r include=FALSE}
  #load libraries
  library(RPostgreSQL)
  library(ggplot2)
  library(lubridate)
  library(scales)
```

## Connect to the database first
```{r}
#loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")
#creates a connection to the postgres database
#note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "twitter",
                 host = "localhost", port = 5432,
                 user = "postgres", password = "")
```

Connection success: `r dbExistsTable(con, c("main","experiment_tweets_shortest"))`

##Pull data from the database
We will only pull the data for one user to understand the structure and content initially

```{r cache=TRUE}
tweets <- dbGetQuery(con, "SELECT \"ID\", \"CREATEDAT\", \"USERID\", \"RETWEET\", \"GEO_ENABLED\", \"LATITUDE\", \"LONGITUDE\", \"LOCATION\", \"TIMEZONE\" from main.experiment_tweets_shortest")
```

Total amount of tweets in the corpus: `r total_tweets <- nrow(tweets)
  total_tweets`

Total unique users: `r total_users <- length(unique(tweets$USERID))
  total_users`

##Dates of the tweets

####Per year
Show a plot off all tweets as per their creation date
```{r}
tweets$CREATEDAT <- ymd_hms(tweets$CREATEDAT)
tweets$YEAR <- year(tweets$CREATEDAT)

ggplot(data = tweets, aes(x = CREATEDAT)) +
  geom_histogram(aes(fill = ..count..)) +
  theme(legend.position = "none") +
  xlab("Time") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

####Per hour
Show a plot off all tweets as per hour
```{r}
ggplot(data = tweets, aes(x = hour(CREATEDAT))) +
  geom_histogram(aes(fill = ..count..)) +
  theme(legend.position = "none") +
  xlab("Hour") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

Is this the same per year?
```{r}
ggplot(data = tweets, aes(x = hour(CREATEDAT))) +
  geom_histogram(aes(fill = ..count..)) +
  theme(legend.position = "none") +
  facet_wrap(~YEAR) +
  xlab("Hour") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

Is the behavior the same for all timezones?

First show the top20 timezones
```{r}
toptz <- data.frame(sort(table(tweets$TIMEZONE),decreasing=T))
toptz <- toptz[1:20,]
ggplot(data = toptz, aes(Var1, Freq)) +
  geom_bar(stat = "identity") +
  theme(legend.position = "none") +
  xlab("Timezone") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Now show the tweets per hour per the top 20 timezones
```{r}
toptz_tweets <- tweets[tweets$TIMEZONE %in% toptz$Var1,]

ggplot(data = toptz_tweets, aes(x = hour(CREATEDAT))) +
  geom_histogram(aes(fill = ..count..)) +
  theme(legend.position = "none") +
  facet_wrap(~TIMEZONE) +
  xlab("Hour") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

####Per weekday
Show a plot off all tweets as per week day
```{r}
ggplot(data = tweets, aes(x = wday(CREATEDAT, label = TRUE))) +
  geom_bar(aes(fill = ..count..)) +
  theme(legend.position = "none") +
  xlab("Day of the Week") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

Is this the same per year?
```{r}
ggplot(data = tweets, aes(x = wday(CREATEDAT, label = TRUE))) +
  geom_bar(aes(fill = ..count..)) +
  theme(legend.position = "none") +
  facet_wrap(~YEAR) +
  xlab("Day of the Week") + ylab("Number of tweets") + 
  scale_fill_gradient(low = "midnightblue", high = "aquamarine4")
```

####Per hour, per location
