---
title: "Tweet Content Information"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
```

```{r include=FALSE}
  #load libraries
  library(knitr)
  library(ggplot2)
```

## Connect to the database first
```{r echo=FALSE}
read_chunk('../DBConnection/ConnectPostgres.R')
```

```{r connectDB}
```

##Pull data from the database
We will only pull the data for one user to understand the structure and content initially

```{r tweets_single_user}
```

```{r}
tweets <- tweets.singleuser
```

Amount of records retrieved: `r total_tweets <- nrow(tweets)
  total_tweets`

## Show first 5 records
```{r}
head(tweets)
```

## Structure of the data

```{r}
str(tweets)
```

## Summary of the data

```{r}
summary(tweets)
```

## Find NA (not applicables) in dataset 
These could be a problem later in data analysis
```{r}
table(is.na(tweets))
```

Are there any complete rows
Incomplete = `r incomplete_tweets <- nrow(tweets[!complete.cases(tweets),])
  incomplete_tweets`

```{r}
df <- data.frame(dose=c("Total", "Incomplete"),len=c(total_tweets, incomplete_tweets))
ggplot(data=df, aes(x=dose, y=len)) + geom_bar(stat="identity", fill="steelblue") + theme_minimal() + coord_flip() + theme(axis.title=element_blank())
#counts = data.frame(total_tweets, incomplete_tweets)
#barplot(as.matrix(counts), main="Total vs Incompleted Rows", horiz=TRUE, names.arg=c("Total", "Incomplete"), row=c("darkblue","red"))
```

Lets see which columns contribute
```{r}
toplot_noind <- data.frame(cols=colnames(tweets), count=colSums(is.na(tweets)))
#remove rows where count = 0
toplot_noind <- toplot_noind[toplot_noind$count > 0,]

ggplot(data=toplot_noind, aes(x = cols, y = count)) + geom_bar(stat="identity", fill="steelblue") + theme_minimal() + theme(axis.title=element_blank())+theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

##Close the Database connection
```{r closeDB}
```