---
title: "K-means clustering"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, messages=FALSE, results='hide'}
#garbage collection + clear RAM
rm(list = ls(all.name = TRUE))
gc()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include=FALSE}
  #load libraries
  library(knitr)
  library(ggplot2)
  library(scales)
```

The K-means algorithm aims to choose centroids C that minimize the within cluster sum of squares objective function with a dataset X with n samples

## Connect to the database first
```{r echo=FALSE}
read_chunk('../DBConnection/ConnectPostgres.R')
```

```{r connectDB}
```

##Get the tweets
```{r gettwitterfeeddata, cache=FALSE}
users <- dbGetQuery(con, "SELECT * from main.experiment_user_shortest")
```

Total amount of users in the corpus: `r users.total <- nrow(users)
  users.total`

```{r preview_users}
#Preview user dataset  
head(users)

#scale the dataset
df <- scale(users[,3:10])

#Preview scaled dataset
head(df)
```

##Train the data
We need to find the optimal amount of clusters first

###Lets look at: no_of_tweets vs no_of_replies:

```{r cluster_size_1}
mydata <- df[, 1:2]

set.seed(20)
wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mydata,centers=i)$withinss)
#par(mar=c(5.1,4.1,4.1,2.1))
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares", main="Within cluster sum of squares (WCSS)")

myCluster <- kmeans(mydata, 8)
users$cluster <- as.factor(myCluster$cluster)
ggplot(users, aes(no_of_tweets, no_of_replies, color = users$cluster)) + geom_point()
```

###Lets look at: no_of_tweets vs no_of_friends:

```{r cluster_size_2}
mydata <- df[, c(1,4)]

set.seed(20)
wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mydata,centers=i)$withinss)
#par(mar=c(5.1,4.1,4.1,2.1))
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares", main="Within cluster sum of squares (WCSS)")

myCluster <- kmeans(mydata, 10)
users$cluster <- as.factor(myCluster$cluster)
ggplot(users, aes(no_of_tweets, no_of_friends, color = users$cluster)) + geom_point()
```

###Lets look at: no_of_friends vs no_of_followers:

```{r cluster_size_3}
mydata <- df[, 4:5]

set.seed(20)
wss <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(mydata,centers=i)$withinss)
#par(mar=c(5.1,4.1,4.1,2.1))
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares", main="Within cluster sum of squares (WCSS)")

myCluster <- kmeans(mydata, 10)
users$cluster <- as.factor(myCluster$cluster)
ggplot(users, aes(no_of_friends, no_of_followers, color = users$cluster)) + geom_point()
```

```{r closeDB}
```