---
title: "Model based clustering"
output: 
  html_document: 
    keep_md: yes
---

```{r echo=FALSE, messages=FALSE, results='hide'}
#garbage collection + clear RAM
rm(list = ls(all.name = TRUE))
gc()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r include=FALSE}
  #load libraries
  library(knitr)
  library(ggplot2)
  library(scales)
```

Model based approaches assume a variety of data models and apply maximum likelihood estimation and Bayes criteria to identify the most likely model and number of clusters. Specifically, the Mclust( ) function in the mclust package selects the optimal model according to BIC for EM initialized by hierarchical clustering for parameterized Gaussian mixture models.

## Connect to the database first
```{r echo=FALSE}
read_chunk('../DBConnection/ConnectPostgres.R')
```

```{r connectDB}
```

##Get the tweets
```{r gettwitterfeeddata, cache=FALSE}
users <- dbGetQuery(con, "SELECT * from main.experiment_user_shortest")
```

Total amount of users in the corpus: `r users.total <- nrow(users)
  users.total`

```{r preview_users}
#Preview user dataset  
head(users)

#scale the dataset
df <- scale(users[,3:10])

#Preview scaled dataset
head(df)
```
  
##Train the data
We need to find the optimal amount of clusters first

###Lets look at: no_of_tweets vs no_of_replies:

```{r cluster_size_1}
mydata <- df

set.seed(20)
library(mclust)

fit <- Mclust(mydata)
plot(fit) # plot results 
summary(fit) # display the best model
```



```{r closeDB}
```